---
title: "Supplementary Material"
format:
    html:
        echo: false
        embed-resources: true
engine: knitr
params:
    counts_file: "results/tables/counts.tsv"
    multiqc_file: "results/multiqc/multiqc_general_stats.txt"
    summary_file: "results/tables/counts.tsv.summary"
    rulegraph_file: "results/rulegraph.png"
    srr_ids: "SRR935090 SRR935091 SRR935092"
    gsm_ids: "GSM1186459 GSM1186460 GSM1186461"
---

```{r Dependencies}
#| include: false
library("ggplot2")
library("reshape2")
library("pheatmap")
library("GEOquery")
```

```{r Read parameters}
#| include: false
counts_file <- params$counts_file
multiqc_file <- params$multiqc_file
summary_file <- params$summary_file
rulegraph_file <- params$rulegraph_file
srr_ids <- unlist(strsplit(params$srr_ids, " "))
gsm_ids <- unlist(strsplit(params$gsm_ids, " "))
```

```{r Read data}
#| include: false
# Read counts
counts <- read.delim(counts_file,
                     header       = TRUE,
                     comment.char = "#",
                     row.names    = 1)
counts <- counts[, 5:ncol(counts)]
colnames(counts) <- gsub(".*(SRR[0-9]+)\\..*", "\\1", colnames(counts))

# Read count summary file
counts_summary <- read.delim(summary_file,
                             sep       = "\t",
                             header    = TRUE,
                             row.names = 1)
colnames(counts_summary) <- gsub(".*(SRR[0-9]+)\\..*", "\\1",
                                 colnames(counts_summary))

# Read metadata
meta <- data.frame()
for (GSM in gsm_ids) {
    gsm <- Meta(getGEO(GSM))
    current_meta <- as.data.frame(
        list(
            title=gsm$title, geo_accession=gsm$geo_accession, source_name_ch1=gsm$source_name_ch1, 
            growth_phase=gsub("growth phase: ", "", gsm$characteristics_ch1[grep("growth phase", gsm$characteristics_ch1)]),
            treatment=gsub("treatment: ", "", gsm$characteristics_ch1[grep("treatment", gsm$characteristics_ch1)])
        )
    )
    meta <- rbind(meta, current_meta)
}
meta <- meta[c("title", "geo_accession", "source_name_ch1", "growth_phase", "treatment")]
gsm2srr <- data.frame(geo_accession = gsm_ids, SRR = srr_ids)
meta <- merge(meta, gsm2srr, by = "geo_accession")

# Read FastQC data and update column names
qc <- read.delim(multiqc_file)
patterns <- c("*.+percent_duplicates.*",
              "*.+percent_gc.*",
              "*.+avg_sequence_length.*",
              "*.+median_sequence_length.*",
              "*.+percent_fails.*",
              "*.+total_sequences.*")
subs <- c("Percent duplicates", "Percent GC", "Avg sequence length",
          "Median sequence length", "Percent fails", "Total sequences")
for (i in 1:length(patterns)) {
    colnames(qc) <- gsub(patterns[i], subs[i], colnames(qc))
}
meta <- merge(meta, qc, by.x = "SRR", by.y = "Sample")
```

# Supplementary Methods

Here you might include a description of the methods used in obtaining and
analysing the data before getting into this document, *e.g.* how sequencing data
was aligned and counted.

# Supplementary Tables and Figures

```{r Sample info}
columns <- c("SRR", "geo_accession", "source_name_ch1", "growth_phase", "treatment")
sample_info <- meta[, columns]
sample_info
```

```{r QC statistics}
columns <- c("SRR", "Percent duplicates", "Percent GC", "Avg sequence length",
             "Percent fails", "Total sequences")
qc_data <- meta[, columns]
qc_data
```

```{r Counts barplot}
count_data <- rbind(genes = apply(counts[, -1], 2, sum), counts_summary)
count_data <- melt(as.matrix(count_data),
                   varnames   = c("Feature", "Sample"),
                   value.name = "Reads")
ggplot(count_data, aes(x = Sample, y = Reads, fill = Feature)) +
    geom_bar(stat = "identity")
```

```{r Gene heatmap}
#| fig-height: 14
cv_cutoff <- 1.2
max_cutoff <- 5
heatmap_data <-
    counts[apply(counts[, -1], 1, function(x) sd(x) / mean(x)) > cv_cutoff &
           apply(counts[, -1], 1, max) > max_cutoff, ]
colnames(heatmap_data)[2:4] <- as.character(meta$title)
labels_row <- paste0(rownames(heatmap_data), " (",
                     substr(gsub("%2C", "", heatmap_data$description), 1, 50),
                     ")")
gg <- pheatmap(log10(heatmap_data[, -1] + 1),
               fontsize_row = 8,
               labels_row = labels_row)
show(gg)
```

## Reproducibility

The code for reproducing this analysis is available in this [GitHub repo](https://github.com/NBISweden/workshop-reproducible-research/tree/master/docker).
The repo contains:

- A Snakemake workflow for running all analysis steps
- A Conda environment file for installing all needed dependenciesx
- A Docker file for running the analysis in a well-defined and isolated system

The results in this supplementary were generated in the following R environment:

```{r Session info}
sessionInfo()
```
